---
title: "Meetup Report"
author: "Greg Sutcliffe"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 9
    fig_height: 4
    fig_caption: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

options(browser="firefox")
library(tidyverse)
library(glue)
library(gt)
library(pins)
library(patchwork)
library(httr)
```

```{r pins}
board <- pins::board_folder('/srv/docker-pins/meetup')  

groups <- pin_read(board, "groups")
events <- pin_read(board, "events") |>
  tidyr::drop_na(id) |>
  left_join(select(groups,urlname,country), by = 'urlname') |>
  mutate(is_online_event = venue_name == 'Online event',
         Location = country)
```

# Meetup - Weekly Readout: `r format(Sys.Date(), '%d %B, %Y')`

Current data on the Ansible Meetup Groups. Changes to which groups are tracked
can be found
[here](https://github.com/ansible-community/stats-crawler/blob/main/config/meetups.yml)

## Meetup activity

Data over the last 12 months-to-date, for all continents:

```{r activity_plot}
summary <- events %>%
  filter(time >= Sys.Date() - months(12)) %>%
  filter(status == 'PAST') %>%
  group_by(urlname,Location) %>%
  mutate(mean_rsvp = round(mean(going,na.rm = TRUE),0)) %>%
  group_by(urlname,Location,mean_rsvp) %>%
  do(
    model = lm(going ~ time, data = .) %>% broom::tidy(),
    n     = nrow(.),
    v     = sum(.$is_online_event)
  ) %>%
  unnest(cols = c(model,n,v)) %>%
  filter(term == 'time') %>%
  replace_na(list(v = 0, p.value = 1)) %>%
  mutate(estimate = estimate*60*60*24, #rsvps / second -> rsvps / day
         trend = case_when(
           p.value  > 0.3 ~ '&#8605;', # squiggle
           estimate > 0   ~ '&#8599;', # up
           estimate < 0   ~ '&#8600;', # down
           TRUE           ~ '??'  # WTF
         ),
         trend = glue("<span style='font-size:20px;'>{trend}</span>")) %>%
  arrange(-n) %>%
  select(-term,-statistic,-std.error,-estimate,-p.value) %>%
  rename(Name = urlname,
         Events = n, `Virtual Events` = v,
         Mean_RSVP = mean_rsvp, RSVP_Trend = trend) %>%
  ungroup()
  
summary %>%
  head(15) %>%
  mutate(Name = str_remove(Name,'Ansible-')) %>%
  ggplot(aes(x=fct_reorder(Name,-Events),y=Events,fill=Mean_RSVP)) +
  geom_bar(stat='identity') +
#  scale_y_discrete() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "grey", midpoint = mean(summary$Mean_RSVP)) +
  theme(axis.text.x = element_text(angle = 70, hjust=1)) +
  xlab('Meetup Name') +
  ylab('Events in the last year')
```

```{r activity_table}
summary %>%
  head(15) %>%
  mutate(Name = glue('[{Name}](https://meetup.com/{Name})')) %>%
  gt() %>%
  fmt_markdown(c('Name','RSVP_Trend'))
```

## Global summaries

Table of past meetup totals by geography, last 12 months:

```{r totals_table}
this_month = cut(Sys.Date(),'month') |> as.Date()
start = this_month - months(13)

# Use pivot_wider_spec to ensure both columns exist even if
# the events are all of one type
pivot_spec <- tibble(
  .name = c("Virtual", "Physical"),
  .value = c("n", "n"),
  is_online_event = c(TRUE,FALSE)
)

events |>
  filter(status == 'PAST') |>
  filter(time >= start & time < this_month) |>
  drop_na() |>
  count(Location, is_online_event) |>
  pivot_wider_spec(pivot_spec, values_fill = 0) |>
  do(bind_rows(., data.frame(Location="Total",
                             count(., name = "Physical", wt = Physical),
                             count(., name = "Virtual", wt = Virtual)))) |>
  gt()
```

Trends in "number of events" and "number of RSVPs", last 12 months, split by type:

```{r global_trends}
events |>
  filter(status == 'PAST') |>
  mutate(month = cut(time, 'month') |> as.Date()) |>
  select(month, going, is_online_event) |>
  count(month, is_online_event, wt = going, name = "going") -> rsvps

events |>
  filter(status == 'PAST') |>
  mutate(month = cut(time, 'month') |> as.Date()) |>
  select(month, title, going, is_online_event) |>
  count(month, is_online_event) |>
  left_join(rsvps, by = c("month","is_online_event")) |>
  drop_na() -> df

df |>
  filter(month >= start) |>
  filter(month < this_month) |>
  ggplot(aes(x = month, y = n, colour = is_online_event, size = going)) +
  geom_point() +
  geom_smooth(method='lm', se = F) +
  theme_minimal() +
  labs(title = "Global Meetup Events / month",
       subtitle = "Last 12 months, size = total RSVPs",
       x = "Month", y = "# of events",
       caption = "Source: Meetup API") -> p1
df |>
  filter(month >= start) |>
  filter(month < this_month) |>
  ggplot(aes(x = month, y = going, colour = is_online_event, size = n)) +
  geom_point() +
  geom_smooth(method='lm', se = F) +
  theme_minimal() +
  labs(title = "Global Meetup Signups / month",
       subtitle = "Last 12 months, size = total events",
       x = "Month", y = "# of signups",
       caption = "Source: Meetup API") -> p2
```

```{r global_trends_plots}
p1 + p2 + plot_layout(guides = 'collect')
```
## Upcoming meetups

All the events where `status = upcoming` and the event time is in the next 3
months:

```{r upcoming-data}
# These also need to be posted to Discourse...
# TODO this needs to be handled separately
#source('discourse.R')

events %>%
  distinct(id, .keep_all = T) %>%
  filter(time > Sys.Date() - lubridate::days(1)) %>%
  filter(time <= Sys.Date() + 90) %>%
  filter(status == 'published') %>%
  #push_to_discourse() %>%
  arrange(time) %>%
  transmute(Event = glue('[{title}]({link})'), 
            time = as.Date(time), Location,
            Group = glue('[{urlname}](https://meetup.com/{urlname})'),
            going, is_online_event,
            #result = glue('[Discourse]({result})')
  ) -> up_events
```

```{r upcoming}
up_events %>%
  gt() %>%
  #fmt_markdown(columns = c('Event','Group', 'result')) %>%
  fmt_markdown(columns = c('Event','Group')) %>%
  cols_label(
    time = "Date",
    going = "Going",
    is_online_event = 'Virtual?',
    #result = 'Discourse'
  )
```

## Recent events

All the events where `status = past` and the event time is in the last month:

```{r recent}
events %>%
  distinct(id, .keep_all = T) %>%
  filter(time >= Sys.Date() - months(1)) %>%
  filter(status == 'PAST') %>%
  arrange(desc(time)) %>%
  transmute(Event = glue('[{title}]({link})'),
            time = as.Date(time), Location,
            going, is_online_event,
            Group = glue('[{urlname}](https://meetup.com/{urlname})')) %>%
  gt() %>%
  fmt_markdown(columns = c('Group','Event')) %>%
  cols_label(
    time = "Date",
    going = "Went",
    is_online_event = 'Virtual?'
  )
```

<!-- ## Untracked groups -->

<!-- **Currently broken** -->

<!-- Groups returned from the Meetup API that are not in the config under either -->
<!-- `allowlist` or `ignorelist`. Add a group to either place to remove it from here. -->

<!-- ```{r eval=FALSE} -->
<!-- raw_groups %>% -->
<!--   anti_join(enframe(config$allowlist), by = c('urlname' = 'value')) %>% -->
<!--   mutate( -->
<!--     url  = paste0('https://meetup.com/',urlname), -->
<!--     Name = map2_chr(name,url, ~ toString(htmltools::tags$a(href=.y,.x)))) %>% -->
<!--   select(Name,city,country,next_event, upcoming_event_count) -> untracked -->
<!-- ``` -->

<!-- There are currently 0 groups that we do not track or ignore, -->
<!-- so only the top 10 (by upcoming event count) are shown here. -->

<!-- ```{r eval=FALSE} -->
<!-- untracked %>% -->
<!--   arrange(-upcoming_event_count) %>% -->
<!--   head(10) %>% -->
<!--   gt() %>% -->
<!--   fmt_markdown(columns = 'Name') %>% -->
<!--     cols_label( -->
<!--     city = "City", -->
<!--     country = "Country", -->
<!--     next_event = "Next Event", -->
<!--     upcoming_event_count = '# Upcoming events' -->
<!--   ) -->
<!-- ``` -->

